---
title: "Data prep"
author: "Bas van Remmerden"
date: "`r Sys.Date()`"
output: html_document
---


# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(123)
```


## Loading libraries
Loading the required libraries
```{r warning = F, message = F}
#### Installing/loading packages
required_packages <- c(
  "tidyverse",
  "lubridate",
  "tidymodels",
  "beepr",
  "DescTools",
  "smotefamily"
)

# install.packages(packages_to_install)
for (pkg in required_packages) {
  require(pkg, character.only = TRUE)
}

# Cleaning the environment before analysis
rm(pkg, required_packages)
```

## Loading linking table
```{r}
linking_table <- read.csv("Original datasets/linking_table_cik_gvkey.csv") %>%
  select(cik, gvkey) %>%
  unique()
```


# Compustat data
## Loading the data
```{r}
df.compustat <- read.csv("Original datasets/financial_information.csv")
```

## Preparation
### Creating year variable
```{r}
df.compustat <- df.compustat %>%
  mutate(year = fyear,
         .after = gvkey)
```

### Dealing with NA values
```{r}
# Variables in the data that I will use
         #rect, # Total receivables 2
         #recd, # Doubtful receivables VEEL NA
         #sale, # Sales 12
         #cshi, # Shares issued
         #csho, # Common shares outstanding
         #oancf, # operating cashflow
         #gp, # gross profit or nisa
         #intan, # Intangible assets, total
         #invt, # Inventory, total
         #act, # Current assets, total
         #at, # assets, total 6
         #cogs, # cost of goods sold 41
         #dd1, # Long term debt due in one year 9
         #lct, # Current liabilities, total 5
         #ppent, # plant property equipment 8
         #xsga, # SGA 198
         #txp, # income taxes payable 71
         #dp, # Depreciation 14
         #am # Amortization 65

set_to_0 <- c("rect",
              "recd",
              "sale",
              "cshi",
              "csho",
              "oancf",
              "gp",
              "intan",
              "invt",
              "cogs",
              "dd1",
              "lct",
              "ppent",
              "xsga",
              "txp",
              "dp",
              "am"
              )

df.compustat <- df.compustat %>%
  mutate_at(vars(one_of(set_to_0)), ~replace_na(., 0))
```


### Creating the variables for my model
```{r}
df.compustat <- df.compustat %>%
  mutate(
    # DSRI
    dsri = (rect/sale)/(lag(rect) / lag(sale)),
    
    # GMI   
    gmi = ((lag(sale)-lag(cogs)) / lag(sale))/((sale/cogs)/sale),
    
    # AQI
    aqi = ((1-act+ppent)/at) / ((1-lag(act)+lag(ppent))/lag(at)),
    
    # SGI     
    sgi = sale / lag(sale),
    
    # DEPI
    depi = (lag(dp)/(lag(dp)+lag(ppent))) / (dp/(dp+ppent)),
    
    # SGAI
    sgai = (xsga / sale) / (lag(xsga)/lag(sale)),
    
    # LVGI
    lvgi = ((dd1+lct)/at) / ((lag(dd1)+lag(lct))/lag(at)),
    
    # new_sh_issued
    new_sh_issued = if_else((csho - lag(csho, default = first(csho)) > 0) | cshi > 0, 1, 0),
    
    # int_assets
    int_assets = intan / at,
    
    # inv_method
    #inv_method = , ###############
    
    # inv_growth
    inv_growth = invt / lag(invt),
    
    # manuf_company
    manuf_company =if_else(sic > 2999 & sic < 4000, 1, 0), 
    
    # cf_earn_corr
    #cf_earn_corr = cor(gp, oancf), ######## 
    
    # cf_earn_smooth
    #cf_earn_smooth = , ########
    
    # afda_rec
    afda_rec = recd / rect,
    
    # afda_sales
    afda_sales = recd / sale,
    
    # rec_sales
    rec_sales = rect / sale,
    
    # inv_sales
    inv_sales = invt / sale
    ) 

df.compustat <- df.compustat %>%
  mutate(
    dsri = ifelse(is.na(dsri), 1, dsri),
    gmi = ifelse(is.na(gmi), 1, gmi),
    aqi = ifelse(is.na(aqi), 1, aqi),
    sgi = ifelse(is.na(sgi), 1, sgi),
    depi = ifelse(is.na(depi), 1, depi),
    sgai = ifelse(is.na(sgai), 1, sgai),
    lvgi = ifelse(is.na(lvgi), 1, lvgi),
  )


```

### Selecting relevant variables
```{r}
df.compustat_filt <- df.compustat %>%
  select(gvkey,
         year,
         sic,
         dsri,
         gmi, 
         aqi,
         sgi,
         depi,
         sgai,
         lvgi,
         new_sh_issued,
         #inv_method,
         inv_growth,
         manuf_company,
         #cf_earn_corr,
         #cf_earn_smooth,
         afda_rec,
         afda_sales,
         rec_sales,
         inv_sales,
         at
         
         #rect, # Total receivables 2
         #recd, # Doubtful receivables VEEL NA
         #sale, # Sales 12
         #cshi, # Shares issued
         #csho, # Common shares outstanding
         #oancf, # operating cashflow
         #gp, # gross profit or nisa
         #intan, # Intangible assets, total
         #invt, # Inventory, total
         #act, # Current assets, total
         #at, # assets, total 6
         #cogs, # cost of goods sold 41
         #dd1, # Long term debt due in one year 9
         #lct, # Current liabilities, total 5
         #ppent, # plant property equipment 8
         #xsga, # SGA 198
         #txp, # income taxes payable 71
         #dp, # Depreciation 14
         #am # Amortization 65
         )
```

### Dropping NA values
```{r}
df.compustat <- df.compustat_filt %>%
  drop_na()
```

### Removing unnecessary objects from my environment
```{r}
rm(df.compustat_filt, set_to_0)
```

### Creating analysis data
```{r}
df.analysis <- df.compustat
```

# Audit fees data
## Loading the data
```{r}
df.audit <- read.csv("Original datasets/audit_fees.csv")
```

## Preparation
### Adding a year denotation
```{r}
df.audit <- df.audit %>%
  mutate(year = year(dmy(FILE_DATE)))
```

### Selecting only necessary variables
```{r}
df.audit <- df.audit %>%
  select(year,
         COMPANY_KEY,
         AUDIT_FEES,
         AUDITOR_FKEY
         ) %>%
  arrange(COMPANY_KEY)
```

### Adding gvkey
```{r}
# Renaming id column in audit data
df.audit <- df.audit %>%
  rename(cik = COMPANY_KEY)

# Linking gvkey to audit
df.audit <- left_join(df.audit, linking_table, by = c("cik"), 
                     relationship = "many-to-many")
```

### Creating big4 indicator
```{r}
df.audit <- df.audit %>%
  mutate(AUDITOR_BIG4 = ifelse(AUDITOR_FKEY %in% c(1, 2, 3, 4), 1, 0))
```


### Linking to analysis data
```{r}
df.analysis <- left_join(df.analysis, df.audit, by = c("gvkey", "year"))  %>%
  select(-cik)
```

### Scaling audit fees
```{r}
df.analysis <- df.analysis %>%
  mutate(FEES_AT = AUDIT_FEES / at) %>%
  select(-at)
```


# Auditor change data
## Preparation
### Loading the data
```{r}
df.audit_change <- read.csv("Original datasets/audit_change.csv")
```

### Adding a year denotation
```{r}
df.audit_change <- df.audit_change %>%
  mutate(year = year(dmy(DISMISS_DATE)))
```

### Selecting only necessary variables
```{r}
df.audit_change <- df.audit_change %>%
  select(year,
         COMPANY_FKEY,
         AUDITOR_RESIGNED
         )
```

### Adding gvkey
```{r}
# Renaming id column in audit data
df.audit_change <- df.audit_change %>%
  rename(cik = COMPANY_FKEY)

# Linking gvkey to audit
df.audit_change <- left_join(df.audit_change, linking_table, by = c("cik"), 
                     relationship = "many-to-many")
```

### Linking to analysis data
```{r}
df.analysis <- left_join(df.analysis, df.audit_change, by = c("gvkey", "year"),
                         relationship = "many-to-many") %>%
  select(-cik)
```

### Replacing NA with 0
```{r}
df.analysis <- df.analysis %>%
  mutate(AUDITOR_RESIGNED = ifelse(is.na(AUDITOR_RESIGNED), 0, AUDITOR_RESIGNED))
```



# Governance data
## Loading the data
```{r}
df.ceo <- read.csv("Original datasets/ceo_data.csv")
```

## Preparation
### Adding gvkey
```{r}
# Renaming id column in ceo data
df.ceo <- df.ceo %>%
  rename(cik = COMPANY_KEY)

# Linking gvkey to audit
df.ceo <- left_join(df.ceo, linking_table, by = c("cik"), 
                     relationship = "many-to-many")
```


### Creating variables
```{r}
# Adding year variable
df.ceo <- df.ceo %>%
  mutate(year = year(dmy(EFF_DATE)))

# Adding variable denoting whether the CEO is also the chairman
df.ceo <- df.ceo %>%
  mutate(ceo_chair = ifelse(IS_CHAIR + IS_CEO == 1, 1, 0))

```
### Selecting relevant variables
```{r}
# Taking unique values to remove duplicates when there are two changes
df.ceo = df.ceo %>%
  filter(ceo_chair == 1) %>%
  select(gvkey, cik, year, ceo_chair) %>%
  unique()
```

### Linking to Analysis data
```{r}
df.analysis <- left_join(df.analysis, df.ceo, by = c("gvkey", "year")) %>%
  select(-cik)
```

### Replacing NA with 0
```{r}
df.analysis <- df.analysis %>%
  mutate(ceo_chair = ifelse(is.na(ceo_chair), 0, ceo_chair))
```


# Financial restatements data
## Preparation
### Loading the data
```{r}
df.rest <- read.csv('Original datasets/financial_restatements.csv')
```

### Setting date variable
Setting the correct format dates
```{r}
df.rest <- df.rest %>%
  mutate(RES_BEGIN_DATE = dmy(RES_BEGIN_DATE),
         RES_END_DATE = dmy(RES_END_DATE)) %>%
  filter(!is.na(RES_BEGIN_DATE) & !is.na(RES_END_DATE))
```

### Adding dependent variable
I follow Karpoff's approach to labeling restatements as irregularities
```{r}
df.rest <- df.rest %>%
  mutate(irregularity = if_else(RES_FRAUD == 1 | 
                                  RES_SEC_INVESTIGATION == 1 | 
                                  (RES_ACCOUNTING == 0 & RES_CLERICAL_ERRORS == 0),
                                1, 0)) 
```

### Adding gvkey
```{r}
# Adding the gvkey through a linking table
df.rest <- df.rest %>%
  mutate(cik = COMPANY_FKEY,
         .before = RESTATEMENT_NOTIFICATION_KEY) %>%
  select(-COMPANY_FKEY,
         -COMPANY_KEY)

# Linking the gvkey to cik
df.rest <- left_join(df.rest, linking_table, by = c("cik"), 
                     relationship = "many-to-many")


```

### Creating a variable denoting auditor change and resignation
```{r}
df.rest <- df.rest %>%
  mutate(AUDITOR_RESIGNED_RES = ifelse(RES_END_AUDITOR_FKEY == CURR_AUD_FKEY, 0, 1))
```

### Changing to long format
By doing this, I can make sure that each year has one observation.
```{r}
df.rest_long <- df.rest %>%
  rowwise() %>%
  do({
    num_years <- year(.$RES_END_DATE) - year(.$RES_BEGIN_DATE) + 1
    data.frame(., year = seq(year(.$RES_BEGIN_DATE), by = 1, length.out = num_years))
  }) %>%
  ungroup()
```

### Saving a copy as the details dataset
```{r}
df.details <- df.rest
df.details_long <- df.rest_long

save(file = "Prepared datasets/details.RData", df.details)
save(file = "Prepared datasets/details_long.RDATA", df.details_long)
```

### Saving relevant variables for analysis data
```{r}
df.rest <- df.rest_long %>%
  select(gvkey,
         year,
         irregularity,
         RES_ADVERSE,
         RES_IMPROVES,
         AUDITOR_RESIGNED_RES
         )

rm(df.rest_long)
```

### Linking to the Analysis data
```{r}
df.analysis <- left_join(df.analysis, df.rest, by = c("gvkey", "year"))
```

### Dealing with NA values
```{r}
df.analysis <- df.analysis %>%
  mutate(RES_ADVERSE = ifelse(is.na(RES_ADVERSE), 2, RES_ADVERSE),
         RES_IMPROVES = ifelse(is.na(RES_IMPROVES), 2, RES_IMPROVES),
         irregularity = ifelse(is.na(irregularity), 2, irregularity),
         
         AUDITOR_RESIGN = ifelse(!is.na(AUDITOR_RESIGNED) & AUDITOR_RESIGNED == 1 | !is.na(AUDITOR_RESIGNED_RES) & AUDITOR_RESIGNED_RES == 1, 1, 0)) %>%
  select(-AUDITOR_RESIGNED,
         -AUDITOR_RESIGNED_RES)
```



# Financial ratios data (v12-05)
## Loading the data
```{r}
df.ratios <- read.csv("Original datasets/financial_ratios.csv") 

```

## Preparation
### Filtering the data
I only want to keep financial ratios for the companies in my restatements data
```{r}
# Obtaining gvkeys from analysis data
company_codes <- unique(df.analysis$gvkey)

# keeping the gvkeys needed
df.ratios <- df.ratios %>%
  filter(gvkey %in% company_codes)
```

### Adding year variable
```{r}
df.ratios <- df.ratios %>%
  mutate(year = ifelse(adate == "", year(dmy(qdate)), year(dmy(adate))),
         .after = "adate")
```

### Changing divyield variable
```{r}
df.ratios <- df.ratios %>%
  mutate(divyield = as.numeric(sub("%", "", divyield))/100,
         year = as.integer(year))
```

### Selecting correct variables
```{r}
df.ratios_filt <- df.ratios %>%
  select(gvkey,
         year, 
         
         # Main financial ratios
         accrual,
         de_ratio,
         bm,
         at_turn,
         npm,
         rect_act,
         
         # Supplementary financial ratios
         inv_turn,
         pay_turn,
         rect_turn,
         invt_act,
         fcf_ocf,
         ocf_lct,
         cash_conversion,
         cash_ratio,
         curr_ratio,
         quick_ratio,
         rd_sale,
         GProf,
         roa,
         roce,
         roe,
         debt_assets,
         pcf,
         pe_inc,
         ps)
```

### Aggregating the data
I have to aggregate the data to obtain yearly ratios instead of monthly ratios. I save the file so that I do not have to run these computations every time.
```{r}
#df.ratios_aggregated <- df.ratios_filt %>%
#  group_by(gvkey, year) %>%
#  summarize(across(where(is.numeric), mean, na.rm = TRUE) , .groups = "keep") 
#  (across(where(is.numeric), ~mean(.x, na.rm = TRUE)), .groups = 'drop')
#beepr::beep(sound = 2)
```

### Saving or loading the data
Either save the data if there are new changes (version should also change) or Load the data
```{r}
# save(file = "Prepared datasets/ratios_long.RData", df.ratios_aggregated)

load(file = "Prepared datasets/ratios_long.RData")
```

### Linking to the analysis data
```{r}
df.analysis_t <- left_join(df.analysis, df.ratios_aggregated, by = c("gvkey", "year"))

df.analysis_t <- drop_na(df.analysis_t)

df.analysis <- df.analysis_t


```

# Analysis data
## Winsorizing values
```{r}
df.analysis <- df.analysis %>%
  mutate(year = as.integer(year),
         AUDIT_FEES = as.numeric(AUDIT_FEES),
         AUDITOR_RESIGN = as.integer(AUDITOR_RESIGN),
         ceo_chair = as.integer(ceo_chair),
         irregularity = as.integer(irregularity),
         RES_ADVERSE = as.integer(RES_ADVERSE),
         RES_IMPROVES = as.integer(RES_IMPROVES),
         new_sh_issued = as.integer(new_sh_issued),
         manuf_company = as.integer(manuf_company)) 

df.analysis <- df.analysis %>%
  mutate(across(where(is.numeric), ~ Winsorize(.x, probs = c(0.01, 0.99))))
```

## Adding additional variables
```{r}
df.analysis <- df.analysis %>%
  mutate(irregularity_bin = ifelse(irregularity == 2, 0, irregularity))
```

## Running SMOTE algorithm
```{r}
df.analysis_filt <- df.analysis %>%
  filter(irregularity != 2)



smote_analysis <- SMOTE(df.analysis_filt[, -ncol(df.analysis_filt)], df.analysis_filt$irregularity, K = 5, dup_size = 0)


df.analysis_smote <- smote_analysis$data

df.analysis_smote <- df.analysis_smote %>% rename(irregularity_smote = class)
```



## Saving the final data
```{r}
save(df.analysis_smote, file = "Prepared datasets/analysis_smote.RData")
save(df.analysis, file = "Prepared datasets/analysis.RData")
```

